---
title: "IMDb Video Games Data Analysis"
author: "L. N."
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r include = FALSE}
getwd()
setwd('C:/Users/17372/Desktop')
games <- read.csv('imdb-videogames.csv')
library(ggplot2)
library(rmarkdown)
```

## Introduction
Thank you for viewing my personal project. This project is still ongoing as I am exploring different ways to visualize and display data to a general audience. This project is programmed in R and the following packages are used: ggplot2, rmarkdown. The data set from this project was gathered from IMDb and collectively published on Kaggle by Muhammad Adil Talay. For more information on the data set, visit <https://www.kaggle.com/datasets/muhammadadiltalay/imdb-video-games>. The following questions I am currently exploring are: What is the average rating for each video game genre? What is the most popular video game in each genre? With no further ado, let's begin by analyzing our first game genre—Action.

## Action Video Games

```{r echo = FALSE, fig.cap = 'Figure 1: Histogram distribution of IMDb ratings for action video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Action == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#AFEEEE', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Action == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Action == 'True'])) + sd(na.omit(games$rating[games$Action == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Action == 'True'])) - sd(na.omit(games$rating[games$Action == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Action Video Games', x = 'Rating', y = 'Action Video Games') + 
  theme_classic()
```

There are `r sum(games$Action == 'True')` action games. The distribution of ratings for action games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Action == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Action == 'True'])), 2)`. The most popular action video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Action == 'True']) == max(na.omit(games$rating[games$Action == 'True'])))]`.

## Adventure Video Games

```{r echo = FALSE, fig.cap = 'Figure 2: Histogram distribution of IMDb ratings for adventure video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Adventure == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#e6e6fa', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Adventure == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Adventure == 'True'])) + sd(na.omit(games$rating[games$Adventure == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Adventure == 'True'])) - sd(na.omit(games$rating[games$Adventure == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Adventure Video Games', x = 'Rating', y = 'Adventure Video Games') + 
  theme_classic()
```

There are `r sum(games$Adventure == 'True')` adventure games. The distribution of ratings for adventure games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Adventure == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Adventure == 'True'])), 2)`. The most popular adventure video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Adventure == 'True']) == max(na.omit(games$rating[games$Adventure == 'True'])))]`.

## Comedy Video Games

```{r echo = FALSE, fig.cap = 'Figure 3: Histogram distribution of IMDb ratings for comedy video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Comedy == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#d3ffce', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Comedy == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Comedy == 'True'])) + sd(na.omit(games$rating[games$Comedy == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Comedy == 'True'])) - sd(na.omit(games$rating[games$Comedy == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Comedy Video Games', x = 'Rating', y = 'Comedy Video Games') + 
  theme_classic()
```

There are `r sum(games$Comedy == 'True')` comedy games. The distribution of ratings for comedy games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Comedy == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Comedy == 'True'])), 2)`. The most popular comedy video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Comedy == 'True']) == max(na.omit(games$rating[games$Comedy == 'True'])))]`.

## Crime Video Games

```{r echo = FALSE, fig.cap = 'Figure 4: Histogram distribution of IMDb ratings for crime video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Crime == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#f08080', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Crime == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Crime == 'True'])) + sd(na.omit(games$rating[games$Crime == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Crime == 'True'])) - sd(na.omit(games$rating[games$Crime == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Crime Video Games', x = 'Rating', y = 'Crime Video Games') + 
  theme_classic()
```

There are `r sum(games$Crime == 'True')` crime games. The distribution of ratings for crime games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Crime == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Crime == 'True'])), 2)`. The most popular crime video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Crime == 'True']) == max(na.omit(games$rating[games$Crime == 'True'])))]`.

## Family Video Games

```{r echo = FALSE, fig.cap = 'Figure 5: Histogram distribution of IMDb ratings for family video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Family == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#ffc3a0', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Family == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Family == 'True'])) + sd(na.omit(games$rating[games$Family == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Family == 'True'])) - sd(na.omit(games$rating[games$Family == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Family Video Games', x = 'Rating', y = 'Family Video Games') + 
  theme_classic()
```

There are `r sum(games$Family == 'True')` family games. The distribution of ratings for family games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Family == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Family == 'True'])), 2)`. The most popular family video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Family == 'True']) == max(na.omit(games$rating[games$Family == 'True'])))]`.

## Fantasy Video Games

```{r echo = FALSE, fig.cap = 'Figure 6: Histogram distribution of IMDb ratings for fantasy video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Fantasy == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#c39797', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Fantasy == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Fantasy == 'True'])) + sd(na.omit(games$rating[games$Fantasy == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Fantasy == 'True'])) - sd(na.omit(games$rating[games$Fantasy == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Fantasy Video Games', x = 'Rating', y = 'Fantasy Video Games') + 
  theme_classic()
```

There are `r sum(games$Fantasy == 'True')` fantasy games. The distribution of ratings for fantasy games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Fantasy == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Fantasy == 'True'])), 2)`. The most popular fantasy video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Fantasy == 'True']) == max(na.omit(games$rating[games$Fantasy == 'True'])))]`.

## Mystery Video Games

```{r echo = FALSE, fig.cap = 'Figure 7: Histogram distribution of IMDb ratings for mystery video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Mystery == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#CBBEB5', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Mystery == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Mystery == 'True'])) + sd(na.omit(games$rating[games$Mystery == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Mystery == 'True'])) - sd(na.omit(games$rating[games$Mystery == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Mystery Video Games', x = 'Rating', y = 'Mystery Video Games') + 
  theme_classic()
```

There are `r sum(games$Mystery == 'True')` mystery games. The distribution of ratings for mystery games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Mystery == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Mystery == 'True'])), 2)`. The most popular mystery video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Mystery == 'True']) == max(na.omit(games$rating[games$Mystery == 'True'])))]`.

## Sci-Fi Video Games

```{r echo = FALSE, fig.cap = 'Figure 8: Histogram distribution of IMDb ratings for sci-fi video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Sci.Fi == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#ff7f50', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Sci.Fi == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Sci.Fi == 'True'])) + sd(na.omit(games$rating[games$Sci.Fi == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Sci.Fi == 'True'])) - sd(na.omit(games$rating[games$Sci.Fi == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Sci-Fi Video Games', x = 'Rating', y = 'Sci-Fi Video Games') + 
  theme_classic()
```

There are `r sum(games$Sci.Fi == 'True')` sci-fi games. The distribution of ratings for sci-fi games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Sci.Fi == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Sci.Fi == 'True'])), 2)`. The most popular sci-fi video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Sci.Fi == 'True']) == max(na.omit(games$rating[games$Sci.Fi == 'True'])))]`.

## Thriller Video Games

```{r echo = FALSE, fig.cap = 'Figure 9: Histogram distribution of IMDb ratings for thriller video games. Ratings range from 0 (min) to 10 (max). The long dashes represent the mean, while the short dashes represent +1 and -1 standard deviation of the mean.'}
ggplot(games[games$Thriller == 'True',]) + 
  geom_histogram(aes(x = rating), fill = '#fa8072', color = 'black', bins = 20) + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Thriller == 'True']))), color = 'black', size = 1, linetype = 'longdash') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Thriller == 'True'])) + sd(na.omit(games$rating[games$Thriller == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  geom_vline(aes(xintercept = mean(na.omit(games$rating[games$Thriller == 'True'])) - sd(na.omit(games$rating[games$Thriller == 'True']))), color = 'black', size = 1, linetype = 'dashed') + 
  labs(title = 'Ratings for Thriller Video Games', x = 'Rating', y = 'Thriller Video Games') + 
  theme_classic()
```

There are `r sum(games$Thriller == 'True')` thriller games. The distribution of ratings for thriller games is left skewed with an average rating of `r round(mean(na.omit(games$rating[games$Thriller == 'True'])), 2)` and a standard deviation of `r round(sd(na.omit(games$rating[games$Thriller == 'True'])), 2)`. The most popular thriller video game(s) by rating is/are `r games$name[which(na.omit(games$rating[games$Thriller == 'True']) == max(na.omit(games$rating[games$Thriller == 'True'])))]`.